---
title: "Modeling"
output: html_document
date: "2025-03-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r packages, include=FALSE}
library(tidyverse)
library(dplyr)
library(Stat2Data)
library(ggplot2)
library(skimr)
library(BMS)
library(leaps)
library(car)
library(caret)
library(reshape2)
library(glmnet)
```

# Calling data

```{r}
defenders <- read.csv("defenders.csv")
defenders <- defenders[, !names(defenders) %in% "X"]
```

# Checking for correlation

```{r}
cor_defenders <- defenders[, !names(defenders) %in% "MP"]
cor_defenders <- cor_defenders[, !names(cor_defenders) %in% "Starts"]
```

```{r}
cor_matrix <- cor(cor_defenders)
```

Build data frame for correlated variables

```{r}
cor_threshold <- 0.80
cor_matrix[lower.tri(cor_matrix, diag = TRUE)] <- NA

cor_data <- as.data.frame(as.table(cor_matrix))

cor_data <- cor_data %>%
  filter(!is.na(Freq) & abs(Freq) > cor_threshold) %>%
  arrange(desc(abs(Freq)))

colnames(cor_data) <- c("Variable1", "Variable2", "Correlation")
```

From here, manually choose which correlated variables to remove from defenders

```{r}
# PrgC v. OneThirdCarry
# DefThirdTkl vs. TklInt
# KP v. Crs
# KP vs. PPA
# PrgDistPass v. OneThirdPass
# PPA v. Crs
defenders <- defenders %>% select (-c(GAminPK, GminusxG, LivePass, xGperNine, TotDistCarry, Rec, TotDistPass, Carries, AttTkl, PrgR, DefPen, ExpPM, CrsPA, Dead, xAGperNine, BlockPass, AttThirdTch, MidThirdTch, DefThirdTch, PrgP, DribbleTkl, MidThirdTkl, xA, onxG, CPA, GlsPerNine, AstPerNine, npxGperNine, ShotsPerNine, Dist, npGminusxG, AttThirdTkl, DefThirdTkl, TklPerc, AxAG, SoTPerc, CmpPercShort, CmpPercMed, CmpPercLong, npxGxAGperNine))
```

# Create data frames for modeling

```{r}
defenders_Starts <- defenders %>% select (-c(MP))
defenders_Starts <- defenders_Starts[, c("Starts", setdiff(names(defenders_Starts), "Starts"))]
```

# Checking for VIF

```{r}
mod2 <- lm(Starts ~. , data = defenders_Starts, singular.ok = TRUE)
summary(mod2)
```

```{r}
vif_fun <- function(df) {
  removed_vars <- list()
  iteration <- 1
  
  while (TRUE) {
    vifs <- car::vif(lm(Starts ~ ., data = df))
    if (max(vifs) < 5) { 
      break
    }
    highest <- names(which(vifs == max(vifs)))
    
    removed_vars[[paste0("Iteration_", iteration)]] <- highest
    
    df <- df[, -which(names(df) %in% highest)]
    
    iteration <- iteration + 1
  }
  
  return(list(cleaned_df = df, removed_vars = removed_vars))
}

result <- vif_fun(defenders_Starts)

defenders_Starts <- result$cleaned_df
removed_variables <- result$removed_vars

print(removed_variables)
```

# Uniform

```{r}
uniform = bms(defenders_Starts,burn=50000, iter=100000, g="BRIC", mprior = "uniform", nmodel=2000, mcmc="bd", user.int=F)
```

# Coefficients for basic mod

```{r}
coef(uniform)
```

# Standardized coefficients for basic mod

```{r}
coef(uniform, std.coefs=T, order.by.pip=F, include.constant=T)
```

# Sampling procedure information

```{r}
summary(uniform)
```

Verify posterior expected model size is equal to the sum of PIP's.

```{r}
sum(coef(uniform)[,1])
```


# Top models from basic BMA

```{r}
topmodels.bma(uniform)[,1:3]
```

# Visualize

```{r}
image(uniform)
```

# Posterior model size plot

```{r}
plotModelsize(uniform)
```

# Convergence between analytical and MCMC PMP's

```{r}
plotConv(uniform)
```

```{r}
plotConv(uniform[1:100])
```

# PMP's of top models

```{r}
pmp.bma(uniform)[1:5,]
```

```{r}
colSums(pmp.bma(uniform))
```

# Compare covariates in terms of PIP by analytical and MCMC methods

```{r}
coef(uniform)[1:11,]
```

```{r}
coef(uniform,exact=TRUE)[1:11,]
```

# Binomial

```{r}
binomial = bms(defenders_Starts, burn=50000, iter=100000, g=10000, mprior="fixed", mprior.size=7, nmodel=2000, mcmc="bd", user.int=F)
```


# Coefficients for binomial mod

```{r}
coef(binomial)
```

# Standardized coefficients for binomial mod

```{r}
coef(binomial, std.coefs=T, order.by.pip=F, include.constant=T)
```

# Sampling procedure information

```{r}
summary(binomial)
```

Verify posterior expected model size is equal to the sum of PIP's.

```{r}
sum(coef(binomial)[,1])
```


# Top models from binomial BMA

```{r}
topmodels.bma(binomial)[,1:3]
```

# Visualize

```{r}
image(binomial)
```

# Posterior model size plot

```{r}
plotModelsize(binomial)
```

# Convergence between analytical and MCMC PMP's

```{r}
plotConv(binomial)
```

```{r}
plotConv(binomial[1:100])
```

# PMP's of top models

```{r}
pmp.bma(binomial)[1:5,]
```

```{r}
colSums(pmp.bma(binomial))
```

# Compare covariates in terms of PIP by analytical and MCMC methods

```{r}
coef(binomial)[1:11,]
```

```{r}
coef(binomial,exact=TRUE)[1:11,]
```

# Beta

```{r}
beta = bms(defenders_Starts, burn=50000, iter=100000, g="BRIC", mprior="random", mprior.size=10, nmodel=2000, mcmc="bd", user.int = F)
```

# Coefficients for beta mod

```{r}
coef(beta)
```

# Standardized coefficients for beta mod

```{r}
coef(beta, std.coefs=T, order.by.pip=F, include.constant=T)
```

# Sampling procedure information

```{r}
summary(beta)
```

Verify posterior expected model size is equal to the sum of PIP's.

```{r}
sum(coef(beta)[,1])
```


# Top models from basic BMA

```{r}
topmodels.bma(beta)[,1:3]
```

# Visualize

```{r}
image(beta)
```

# Posterior model size plot

```{r}
plotModelsize(beta)
```

# Convergence between analytical and MCMC PMP's

```{r}
plotConv(beta)
```

```{r}
plotConv(beta[1:100])
```

# PMP's of top models

```{r}
pmp.bma(beta)[1:5,]
```

```{r}
colSums(pmp.bma(beta))
```

# Compare covariates in terms of PIP by analytical and MCMC methods

```{r}
coef(beta)[1:11,]
```

```{r}
coef(beta,exact=TRUE)[1:11,]
```

# Comparing simulations

```{r}
plotComp(Uniform=uniform, Fixed=binomial, Random=beta)
```

# Marginal Density

```{r}
density(uniform,reg="SoTPerNine", addons = 'Eebl')
```

```{r}
density(beta,reg="SoTPerNine", addons = 'Eebl')
```

# Predictive Density

```{r}
fcstbma= bms(defenders_Starts[1:150,], burn=50000, iter=100000, g="BRIC", mprior="random", mprior.size=10, nmodel=2000, mcmc="bd", user.int = F)

pdens = pred.density(fcstbma, newdata=defenders_Starts[151:152,])
```

```{r}
pdens
plot(pdens, 1)
plot(pdens, 2)
```

```{r}
quantile(pdens, c(0.05, 0.95))
```

```{r}
pdens$fit - defenders_Starts[151:152,1]
```

```{r}
plot(pdens, realized.y = defenders_Starts[151:152, "Starts"])
```

```{r}
lps.bma(pdens, defenders_Starts[151:152,1])
```

# Investigating SCANinety

```{r}
defenders_Starts$SCANinety_sq <- defenders_Starts$SCANinety^2
SCA = bms(defenders_Starts, burn=50000, iter=100000, g="UIP", mprior="random", mprior.size=10, nmodel=2000, mcmc="bd", user.int = F)
```

```{r}
coef(SCA)[c("SCANinety", "SCANinety_sq"), ]
```


```{r}
ggplot(defenders_Starts, aes(SCANinety, Starts)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="loess", span=0.8, color="blue")  +
  geom_vline(xintercept = quantile(defenders_Starts$SCANinety, 0.75), 
             linetype="dashed", color="darkgreen")
```

```{r}
high_SCA_cutoff <- quantile(defenders_Starts$SCANinety, 0.75)
defenders_Starts$high_sca <- as.numeric(defenders_Starts$SCANinety >= high_SCA_cutoff)
high_SCA = bms(defenders_Starts, burn=50000, iter=100000, g="UIP", mprior="random", mprior.size=10, nmodel=2000, mcmc="bd", user.int = F)
```

```{r}
coef(high_SCA)[c("SCANinety", "high_sca"), ]
```

# Investigating SoTPerNine

```{r}
defenders_Starts$SoTPerNine_sq <- defenders_Starts$SoTPerNine^2
SoT = bms(defenders_Starts, burn=50000, iter=100000, g="UIP", mprior="random", mprior.size=10, nmodel=2000, mcmc="bd", user.int = F)
```

```{r}
coef(SoT)[c("SoTPerNine", "SoTPerNine_sq"), ]
```


```{r}
ggplot(defenders_Starts, aes(SoTPerNine, Starts)) +
  geom_point(alpha=0.6) +
  geom_smooth(method="gam", formula = y ~ s(x, bs="tp"), color="red")
```


```{r}
high_sot_cutoff <- quantile(defenders_Starts$SoTPerNine, 0.75)
defenders_Starts$high_sot <- as.numeric(defenders_Starts$SoTPerNine >= high_sot_cutoff)
high_SoT = bms(defenders_Starts, burn=50000, iter=100000, g="UIP", mprior="random", mprior.size=10, nmodel=2000, mcmc="bd", user.int = F)
```

```{r}
coef(high_SoT)[c("SoTPerNine", "high_sot"), ]
```

High PIP predictors

```{r}
high_pip_predictors <- c("AttPen", "OneThirdCarry", "OneThirdPass", "Fls", "Fld", "Crs", "Clr", "Blocks", "CmpPercTotal")
defenders_PIP2 <- defenders_Starts[, c("Starts", high_pip_predictors)]
```

Simple Model

```{r}
simple_model2 <- lm(Starts ~ ., data = defenders_PIP2)
summary(simple_model2)
```

Regression conditions

```{r}
plot(simple_model2, which = 1)
plot(simple_model2, which = 2)
ggplot(defenders_Starts, aes(x=residuals(simple_model2))) + geom_histogram()
plot(simple_model2, which = 3)
plot(simple_model2, which = 5)
```


