---
title: "Scraping and Cleaning 2023-2024"
output: html_document
date: "2025-01-29"
---

Packages

```{r packages, include=FALSE}
library(tidyverse,quietly=TRUE)
library(rvest,quietly=TRUE)
library(janitor,quietly=TRUE)
library(dplyr,quietly=TRUE)
library(prismatic,quietly=TRUE)
library(broom,quietly=TRUE)
library(purrr,quietly=TRUE)
```

Scraper

```{r}
urls = c('https://fbref.com/en/comps/Big5/2022-2023/stats/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/shooting/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/passing/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/passing_types/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/defense/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/gca/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/possession/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/playingtime/players/2022-2023-Big-5-European-Leagues-Stats',
         'https://fbref.com/en/comps/Big5/2022-2023/misc/players/2022-2023-Big-5-European-Leagues-Stats')

stats=read_html(urls[1],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

shooting=read_html(urls[2],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

passing=read_html(urls[3],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

passing_types=read_html(urls[4],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

defence=read_html(urls[5],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

gca=read_html(urls[6],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

possession=read_html(urls[7],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

playingtime=read_html(urls[8],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)

misc=read_html(urls[9],as.data.frame=TRUE,stringAsFactors=TRUE) %>% html_nodes("table") %>% .[[1]] %>% html_table(fill=TRUE)
```

Fixing headers

```{r}
new_header <- as.character(defence[1, ])  
defence <- defence[-1, ]  
colnames(defence) <- new_header

new_header <- as.character(gca[1, ])  
gca <- gca[-1, ]  
colnames(gca) <- new_header

new_header <- as.character(misc[1, ])  
misc <- misc[-1, ]  
colnames(misc) <- new_header

new_header <- as.character(passing[1, ])  
passing <- passing[-1, ]  
colnames(passing) <- new_header

new_header <- as.character(passing_types[1, ])
passing_types <- passing_types[-1, ]  
colnames(passing_types) <- new_header

new_header <- as.character(playingtime[1, ])
playingtime <- playingtime[-1, ]
colnames(playingtime) <- new_header

new_header <- as.character(possession[1, ])
possession <- possession[-1, ]
colnames(possession) <- new_header

new_header <- as.character(shooting[1, ]) 
shooting <- shooting[-1, ]
colnames(shooting) <- new_header

new_header <- as.character(stats[1, ])
stats <- stats[-1, ]  
colnames(stats) <- new_header
```

Cleaning intermittent headers out

```{r}
for (i in seq(26, nrow(defence), by = 26)) {
  defence[i, ] <- NA
}
defence <- na.omit(defence)
for (i in seq(26, nrow(gca), by = 26)) {
  gca[i, ] <- NA
}
gca <- na.omit(gca)
for (i in seq(26, nrow(misc), by = 26)) {
  misc[i, ] <- NA
}
misc <- na.omit(misc)
for (i in seq(26, nrow(passing), by = 26)) {
  passing[i, ] <- NA
}
passing <- na.omit(passing)
for (i in seq(26, nrow(passing_types), by = 26)) {
  passing_types[i, ] <- NA
}
passing_types <- na.omit(passing_types)
for (i in seq(26, nrow(playingtime), by = 26)) {
  playingtime[i, ] <- NA
}
playingtime <- na.omit(playingtime)
for (i in seq(26, nrow(possession), by = 26)) {
  possession[i, ] <- NA
}
possession <- na.omit(possession)
for (i in seq(26, nrow(shooting), by = 26)) {
  shooting[i, ] <- NA
}
shooting <- na.omit(shooting)
for (i in seq(26, nrow(stats), by = 26)) {
  stats[i, ] <- NA
}
stats <- na.omit(stats)
```

Cleaning Nation, Squad, Rk, and Matches columns out

```{r}
defence <- defence[, !names(defence) %in% "Nation"]
defence <- defence[, !names(defence) %in% "Squad"]
gca <- gca[, !names(gca) %in% "Nation"]
gca <- gca[, !names(gca) %in% "Squad"]
misc <- misc[, !names(misc) %in% "Nation"]
misc <- misc[, !names(misc) %in% "Squad"]
passing <- passing[, !names(passing) %in% "Nation"]
passing <- passing[, !names(passing) %in% "Squad"]
passing_types <- passing_types[, !names(passing_types) %in% "Nation"]
passing_types <- passing_types[, !names(passing_types) %in% "Squad"]
playingtime <- playingtime[, !names(playingtime) %in% "Nation"]
playingtime <- playingtime[, !names(playingtime) %in% "Squad"]
possession <- possession[, !names(possession) %in% "Nation"]
possession <- possession[, !names(possession) %in% "Squad"]
shooting <- shooting[, !names(shooting) %in% "Nation"]
shooting <- shooting[, !names(shooting) %in% "Squad"]
stats <- stats[, !names(stats) %in% "Nation"]
stats <- stats[, !names(stats) %in% "Squad"]
```

```{r}
defence <- defence[, !names(defence) %in% "Rk"]
defence <- defence[, !names(defence) %in% "Matches"]
gca <- gca[, !names(gca) %in% "Rk"]
gca <- gca[, !names(gca) %in% "Matches"]
misc <- misc[, !names(misc) %in% "Rk"]
misc <- misc[, !names(misc) %in% "Matches"]
passing <- passing[, !names(passing) %in% "Rk"]
passing <- passing[, !names(passing) %in% "Matches"]
passing_types <- passing_types[, !names(passing_types) %in% "Rk"]
passing_types <- passing_types[, !names(passing_types) %in% "Matches"]
playingtime <- playingtime[, !names(playingtime) %in% "Rk"]
playingtime <- playingtime[, !names(playingtime) %in% "Matches"]
possession <- possession[, !names(possession) %in% "Rk"]
possession <- possession[, !names(possession) %in% "Matches"]
shooting <- shooting[, !names(shooting) %in% "Rk"]
shooting <- shooting[, !names(shooting) %in% "Matches"]
stats <- stats[, !names(stats) %in% "Rk"]
stats <- stats[, !names(stats) %in% "Matches"]
```

Clean defence column names

```{r}
names(defence)[names(defence) == "Def 3rd"] <- "DefThird"
names(defence)[names(defence) == "Mid 3rd"] <- "MidThird"
names(defence)[names(defence) == "90s"] <- "Nineties"
names(defence)[names(defence) == "Att 3rd"] <- "AttThird"
names(defence)[names(defence) == "Tkl%"] <- "TklPerc"
names(defence)[names(defence) == "Tkl+Int"] <- "TklInt"
colnames(defence)[which(colnames(defence) == "Tkl")][2] <- "DribbleTkl"
```

Clean GCA column names

```{r}
names(gca)[names(gca) == "90s"] <- "Nineties"
names(gca)[names(gca) == "SCA90"] <- "SCANinety"
names(gca)[names(gca) == "GCA90"] <- "GCANinety"
colnames(gca)[which(colnames(gca) == "PassLive")][1] <- "PassLiveSCA"
colnames(gca)[which(colnames(gca) == "PassDead")][1] <- "PassDeadSCA"
colnames(gca)[which(colnames(gca) == "TO")][1] <- "TOSCA"
colnames(gca)[which(colnames(gca) == "Sh")][1] <- "ShSCA"
colnames(gca)[which(colnames(gca) == "Fld")][1] <- "FldSCA"
colnames(gca)[which(colnames(gca) == "Def")][1] <- "DefSCA"
names(gca)[names(gca) == "PassLive"] <- "PassLiveGCA"
names(gca)[names(gca) == "PassDead"] <- "PassDeadGCA"
names(gca)[names(gca) == "TO"] <- "TOGCA"
names(gca)[names(gca) == "Sh"] <- "ShGCA"
names(gca)[names(gca) == "Fld"] <- "FldGCA"
names(gca)[names(gca) == "Def"] <- "DefGCA"
```

Clean misc column names

```{r}
names(misc)[names(misc) == "90s"] <- "Nineties"
names(misc)[names(misc) == "2CrdY"] <- "TwoCrdY"
names(misc)[names(misc) == "Won%"] <- "WonPerc"
```

Clean passing column names

```{r}
names(passing)[names(passing) == "90s"] <- "Nineties"
colnames(passing)[which(colnames(passing) == "Cmp")][1] <- "CmpTotal"
colnames(passing)[which(colnames(passing) == "Att")][1] <- "AttTotal"
colnames(passing)[which(colnames(passing) == "Cmp%")][1] <- "CmpPercTotal"
colnames(passing)[which(colnames(passing) == "Cmp")][1] <- "CmpShort"
colnames(passing)[which(colnames(passing) == "Att")][1] <- "AttShort"
colnames(passing)[which(colnames(passing) == "Cmp%")][1] <- "CmpPercShort"
colnames(passing)[which(colnames(passing) == "Cmp")][1] <- "CmpMed"
colnames(passing)[which(colnames(passing) == "Att")][1] <- "AttMed"
colnames(passing)[which(colnames(passing) == "Cmp%")][1] <- "CmpPercMed"
names(passing)[names(passing) == "Cmp"] <- "CmpLong"
names(passing)[names(passing) == "Att"] <- "AttLong"
names(passing)[names(passing) == "Cmp%"] <- "CmpPercLong"
names(passing)[names(passing) == "A-xAG"] <- "AxAG"
names(passing)[names(passing) == "1/3"] <- "OneThird"
```

Cleanign passing_types column names

```{r}
names(passing_types)[names(passing_types) == "90s"] <- "Nineties"
```

Clean playingtime column names

```{r}
names(playingtime)[names(playingtime) == "Mn/MP"] <- "MnPerMatch"
names(playingtime)[names(playingtime) == "Min%"] <- "MinPerc"
names(playingtime)[names(playingtime) == "90s"] <- "Nineties"
names(playingtime)[names(playingtime) == "Mn/Start"] <- "MnPerStart"
names(playingtime)[names(playingtime) == "Mn/Sub"] <- "MnPerSub"
names(playingtime)[names(playingtime) == "+/-"] <- "PlusMinus"
names(playingtime)[names(playingtime) == "+/-90"] <- "PMperNine"
names(playingtime)[names(playingtime) == "xG+/-"] <- "ExpPM"
names(playingtime)[names(playingtime) == "xG+/-90"] <- "ExpPMperNine"
colnames(playingtime)[which(colnames(playingtime) == "On-Off")][1] <- "OnOff"
names(playingtime)[names(playingtime) == "On-Off"] <- "ExpOnOff"
```

Cleaning possession column names

```{r}
names(possession)[names(possession) == "90s"] <- "Nineties"
names(possession)[names(possession) == "Def Pen"] <- "DefPen"
names(possession)[names(possession) == "Def 3rd"] <- "DefThird"
names(possession)[names(possession) == "Mid 3rd"] <- "MidThird"
names(possession)[names(possession) == "Att 3rd"] <- "AttThird"
names(possession)[names(possession) == "Att Pen"] <- "AttPen"
names(possession)[names(possession) == "Succ%"] <- "SuccPerc"
names(possession)[names(possession) == "Tkld%"] <- "TkldPerc"
names(possession)[names(possession) == "1/3"] <- "OneThird"
```

Cleaning shooting column names

```{r}
names(shooting)[names(shooting) == "90s"] <- "Nineties"
names(shooting)[names(shooting) == "SoT%"] <- "SoTPerc"
names(shooting)[names(shooting) == "Sh/90"] <- "ShotsPerNine"
names(shooting)[names(shooting) == "SoT/90"] <- "SoTPerNine"
names(shooting)[names(shooting) == "G/Sh"] <- "GPerS"
names(shooting)[names(shooting) == "G/SoT"] <- "GPerSoT"
names(shooting)[names(shooting) == "npxG/Sh"] <- "npxGperSh"
names(shooting)[names(shooting) == "G-xG"] <- "GminusxG"
names(shooting)[names(shooting) == "np:G-xG"] <- "npGminusxG"
```

Cleaning stats column names

```{r}
colnames(stats)[which(colnames(stats) == "Gls")][2] <- "GlsPerNine"
colnames(stats)[which(colnames(stats) == "Ast")][2] <- "AstPerNine"
colnames(stats)[which(colnames(stats) == "G-PK")][2] <- "GPKperNine"
colnames(stats)[which(colnames(stats) == "xG")][2] <- "xGperNine"
colnames(stats)[which(colnames(stats) == "xAG")][2] <- "xAGperNine"
colnames(stats)[which(colnames(stats) == "npxG")][2] <- "npxGperNine"
colnames(stats)[which(colnames(stats) == "npxG+xAG")][2] <- "npxGxAGperNine"
colnames(stats)[which(colnames(stats) == "G+A")][2] <- "GAPerNine"
names(stats)[names(stats) == "90s"] <- "Nineties"
names(stats)[names(stats) == "G+A"] <- "GA"
names(stats)[names(stats) == "G-PK"] <- "GminPK"
names(stats)[names(stats) == "npxG+xAG"] <- "npxGxAG"
names(stats)[names(stats) == "G+A-PK"] <- "GAminPK"
names(stats)[names(stats) == "xG+xAG"] <- "xGxAG"
```

Cleaning out repeating variable across data frames (i.e. Pos, Comp, Age, Born, etc.)

```{r}
defence <- defence[, !names(defence) %in% "Pos"]
defence <- defence[, !names(defence) %in% "Comp"]
gca <- gca[, !names(gca) %in% "Pos"]
gca <- gca[, !names(gca) %in% "Comp"]
misc <- misc[, !names(misc) %in% "Pos"]
misc <- misc[, !names(misc) %in% "Comp"]
passing <- passing[, !names(passing) %in% "Pos"]
passing <- passing[, !names(passing) %in% "Comp"]
passing_types <- passing_types[, !names(passing_types) %in% "Pos"]
passing_types <- passing_types[, !names(passing_types) %in% "Comp"]
playingtime <- playingtime[, !names(playingtime) %in% "Pos"]
playingtime <- playingtime[, !names(playingtime) %in% "Comp"]
possession <- possession[, !names(possession) %in% "Pos"]
possession <- possession[, !names(possession) %in% "Comp"]
shooting <- shooting[, !names(shooting) %in% "Pos"]
shooting <- shooting[, !names(shooting) %in% "Comp"]
defence <- defence[, !names(defence) %in% "Age"]
defence <- defence[, !names(defence) %in% "Born"]
gca <- gca[, !names(gca) %in% "Age"]
gca <- gca[, !names(gca) %in% "Born"]
misc <- misc[, !names(misc) %in% "Age"]
misc <- misc[, !names(misc) %in% "Born"]
passing <- passing[, !names(passing) %in% "Age"]
passing <- passing[, !names(passing) %in% "Born"]
passing_types <- passing_types[, !names(passing_types) %in% "Age"]
passing_types <- passing_types[, !names(passing_types) %in% "Born"]
playingtime <- playingtime[, !names(playingtime) %in% "Age"]
playingtime <- playingtime[, !names(playingtime) %in% "Born"]
possession <- possession[, !names(possession) %in% "Age"]
possession <- possession[, !names(possession) %in% "Born"]
shooting <- shooting[, !names(shooting) %in% "Age"]
shooting <- shooting[, !names(shooting) %in% "Born"]
```

```{r}
defence <- defence[, !names(defence) %in% "Nineties"]
gca <- gca[, !names(gca) %in% "Nineties"]
misc <- misc[, !names(misc) %in% "Nineties"]
passing <- passing[, !names(passing) %in% "Nineties"]
passing_types <- passing_types[, !names(passing_types) %in% "Nineties"]
playingtime <- playingtime[, !names(playingtime) %in% "Nineties"]
possession <- possession[, !names(possession) %in% "Nineties"]
shooting <- shooting[, !names(shooting) %in% "Nineties"]
```

```{r}
playingtime <- playingtime[, !names(playingtime) %in% "MP"]
playingtime <- playingtime[, !names(playingtime) %in% "Min"]
playingtime <- playingtime[, !names(playingtime) %in% "Starts"]
```

```{r}
shooting <- shooting[, !names(shooting) %in% "Gls"]
shooting <- shooting[, !names(shooting) %in% "PK"]
shooting <- shooting[, !names(shooting) %in% "PKatt"]
shooting <- shooting[, !names(shooting) %in% "xG"]
shooting <- shooting[, !names(shooting) %in% "npxG"]
names(shooting)[names(shooting) == 'FK'] <- 'FKShot'
```

```{r}
names(possession)[names(possession) == 'DefThird'] <- 'DefThirdTch'
names(possession)[names(possession) == 'MidThird'] <- 'MidThirdTch'
names(possession)[names(possession) == 'AttThird'] <- 'AttThirdTch'
names(possession)[names(possession) == 'Live'] <- 'LiveTch'
names(possession)[names(possession) == 'Att'] <- 'AttDrib'
names(possession)[names(possession) == 'TotDist'] <- 'TotDistCarry'
names(possession)[names(possession) == 'PrgDist'] <- 'PrgDistCarry'
possession <- possession[, !names(possession) %in% "PrgC"]
names(possession)[names(possession) == 'OneThird'] <- 'OneThirdCarry'
possession <- possession[, !names(possession) %in% "PrgR"]
```

```{r}
names(passing_types)[names(passing_types) == 'Live'] <- 'LivePass'
passing_types <- passing_types[, !names(passing_types) %in% "Att"]
names(passing_types)[names(passing_types) == 'FK'] <- 'FKPass'
passing_types <- passing_types[, !names(passing_types) %in% "Crs"]
passing_types <- passing_types[, !names(passing_types) %in% "Cmp"]
names(passing_types)[names(passing_types) == 'Off'] <- 'OffPass'
names(passing_types)[names(passing_types) == 'Blocks'] <- 'PassBlock'
```

```{r}
names(passing)[names(passing) == 'TotDist'] <- 'TotDistPass'
names(passing)[names(passing) == 'PrgDist'] <- 'PrgDistPass'
passing <- passing[, !names(passing) %in% "Ast"]
passing <- passing[, !names(passing) %in% "xAG"]
names(passing)[names(passing) == 'OneThird'] <- 'OneThirdPass'
passing <- passing[, !names(passing) %in% "PrgP"]
```

```{r}
misc <- misc[, !names(misc) %in% "CrdY"]
misc <- misc[, !names(misc) %in% "CrdR"]
misc <- misc[, !names(misc) %in% "Int"]
misc <- misc[, !names(misc) %in% "TklW"]
```

```{r}
names(defence)[names(defence) == 'DefThird'] <- 'DefThirdTkl'
names(defence)[names(defence) == 'MidThird'] <- 'MidThirdTkl'
names(defence)[names(defence) == 'AttThird'] <- 'AttThirdTkl'
names(defence)[names(defence) == 'Att'] <- 'AttTkl'
names(defence)[names(defence) == 'Lost'] <- 'LostTkl'
names(defence)[names(defence) == 'Sh'] <- 'BlockSh'
names(defence)[names(defence) == 'Pass'] <- 'BlockPass'
```


Recode position column

```{r}
recode_positions <- function(df) {
    df$Pos <- sapply(strsplit(as.character(df$Pos), ","), `[`, 1)
    
    df$Pos[df$Pos == "GK"] <- 0
    df$Pos[df$Pos == "DF"] <- 1
    df$Pos[df$Pos == "MF"] <- 2
    df$Pos[df$Pos == "FW"] <- 3

    df$Pos <- as.numeric(as.character(df$Pos))
    
    return(df)
}

stats <- recode_positions(stats)
```

Recode competitions column

```{r}
recode_competitions <- function(df) {
    df$Comp[df$Comp == "eng Premier League"] <- 0
    df$Comp[df$Comp == "it Serie A"] <- 1
    df$Comp[df$Comp == "es La Liga"] <- 2
    df$Comp[df$Comp == "fr Ligue 1"] <- 3
    df$Comp[df$Comp == "de Bundesliga"] <- 4

    df$Comp <- as.numeric(as.character(df$Comp))
    
    return(df)
}

stats <- recode_competitions(stats)
```

Cleaning players out of playingtime that are not included in the other data frames

```{r}
playingtime <- playingtime[playingtime$Player %in% defence$Player, ]
```

Merge data frames

```{r}
stats <- stats %>%
  left_join(shooting, by = "Player")

stats <- stats %>%
  left_join(possession, by = "Player")

stats <- stats %>%
  left_join(playingtime, by = "Player")

stats <- stats %>%
  left_join(passing_types, by = "Player")

stats <- stats %>%
  left_join(passing, by = "Player")

stats <- stats %>%
  left_join(misc, by = "Player")

stats <- stats %>%
  left_join(gca, by = "Player")

stats <- stats %>%
  left_join(defence, by = "Player")
```

```{r}
merged_df <- stats %>%
  group_by(Player) %>%
  slice(1) %>%
  ungroup()
```


Clean empty cells/rows/entries

```{r}
clean_data_frame <- function(df) {
    df <- df[!rowSums(is.na(df) | df == "") > 0, ]
    return(df)
}

merged_df <- clean_data_frame(merged_df)
```

Write merged_df to .csv format

```{r}
write.csv(merged_df, '~/Desktop/Senior Seminar/merged_df.csv')
```




